---
title: "Project2 Technical Notebook"
output: html_notebook
---
load in tidyverse and  libraries needed for statistical tests
```{r}
library(tidyverse)
library(pwr)
library(car)
library(effectsize)
```


read in the email metrics csv 
```{r}
email_csv <- read_csv("data/email_metrics.csv")

```


read in the last email before order csv 
```{r}
last_email_csv <- read_csv("data/last_email_before_order.csv")
```


we can see that the columns of group and group code are NOT the same thing; group specifies the general category (conversational approach vs direct) whereas group_code has subcategories within each approach. 0 2 and 4 likely correspond to the days of the week (mon, wed, fri) the emails were sent
```{r}
distinct(email_csv, group, group_code)
```


create copy of original email mentrics dataset  to preserve original version
```{r}
cleaned_email_csv <- email_csv 

```


do the same for last_email_sent dataset 
```{r}
cleaned_last_email_csv <- last_email_csv
```


based on findings, due to the column number_ordered not having a single instance of 0 orders placed, we can interpret NA as there being 0 orders 
```{r}
cleaned_email_csv$number_orders[is.na(cleaned_email_csv$number_orders)] <- 0
```


we can remove the ...1 column as it is just the row index 
```{r}
cleaned_email_csv <- cleaned_email_csv %>% 
  select(-`...1`)
```


ensure that there isnt a instance of there being 0 emails sent - there is for number_opens although we can leave this as NaN because 0 opens means the email itself was never interacted with by a user 
```{r}
any(cleaned_email_csv$emails_sent == 0)
any(cleaned_email_csv$number_opens == 0)
```


create calculated field columns that measures open rate, click rate, open to click rate, unsub rate, and order rate 
```{r}
cleaned_email_csv <- cleaned_email_csv %>% 
  mutate(open_rate = (number_opens / emails_sent),
         click_rate = (number_clicks / emails_sent),
         open_to_click_rate = (number_clicks / number_opens),
         unsubscribe_rate = (number_unsub / emails_sent),
         order_rate = (number_orders / emails_sent)
  )
         

```


calculate the mean value based on tone for unsubscribe rate (conversational vs alarming)
```{r}
cleaned_email_csv %>% 
  group_by(group) %>% 
  summarise(mean_val = mean(unsubscribe_rate, na.rm = TRUE))
```


check p value
```{r}
leveneTest(unsubscribe_rate ~ group, data = cleaned_email_csv)
```


our p value is < 0.05 so we assume there is evidence to reject the null hypothesis, run a welch test
```{r}
t.test(unsubscribe_rate ~ group, alternative = "two.sided", conf.level = 0.95, data = cleaned_email_csv)
```


create unsubscription graph
```{r}
unsubscribe_rate_graph <- cleaned_email_csv %>%
  group_by(group) %>%
  summarise(
    mean_unsub_rate = mean(unsubscribe_rate, na.rm = TRUE),
  )

ggplot(unsubscribe_rate_graph, aes(x = group, 
                                   y = mean_unsub_rate,
                                   fill = group)) +
  labs(title = "Average Unsubscribe Rate Based on Tone",
       x = "Tone",
       y = "Unsubscribe Rate",
       fill = "Tone") +
  geom_col() +
  scale_x_discrete(labels = c("Group A" = "Conversational",
                              "Group B" = "Alarming")) +
  scale_fill_discrete(labels = c("Group A" = "Conversational",
                                 "Group B" = "Alarming"))
```


calculate the mean value based on tone for order rate (conversational vs alarming)
```{r}
cleaned_email_csv %>% 
  group_by(group) %>% 
  summarise(mean_val = mean(order_rate, na.rm = TRUE))
```


check p value
```{r}
leveneTest(order_rate ~ group, data = cleaned_email_csv)
```


our p value is < 0.05 so we assume there is evidence to reject the null hypothesis, run a welch test
```{r}
t.test(order_rate ~ group, alternative = "two.sided", conf.level = 0.95, data = cleaned_email_csv)
```


make order graph
```{r}
order_rate_graph <- cleaned_email_csv %>%
  group_by(group) %>%
  summarise(
    mean_order_rate = mean(order_rate, na.rm = TRUE),
  )

ggplot(order_rate_graph, aes(x = group, 
                                   y = mean_order_rate,
                                   fill = group)) +
  labs(title = "Average Order Rate Based on Tone",
       x = "Tone",
       y = "Order Rate",
       fill = "Tone") +
  geom_col() +
  scale_x_discrete(labels = c("Group A" = "Conversational",
                              "Group B" = "Alarming")) +
  scale_fill_discrete(labels = c("Group A" = "Conversational",
                                 "Group B" = "Alarming"))
```


create sales csv
```{r}
sales_data_csv <- read_csv("data/sales_data_sample.csv")
```


make orders data frame merge order numbers
```{r}
orders_df <- merge(cleaned_last_email_csv,
                sales_data_csv[, c("ORDERNUMBER", "SALES")], 
                by = "ORDERNUMBER")
head(orders_df, 10)
```


sum sales per order number include group
```{r}
order_number_totals <- aggregate(SALES ~ ORDERNUMBER + group, 
                                 data = orders_df, 
                                 sum)
head(order_number_totals, 45)
```


sum sales per group type
```{r}
group_totals <- aggregate(SALES ~ group, 
                          data = order_number_totals, 
                          sum)
head(group_totals, 10)
```


test
```{r}
leveneTest(SALES ~ group, data = group_totals)
```


sales amount (conversational vs alarming)
```{r}
options(scipen = 999)
ggplot(group_totals, aes(x = group, y = SALES, fill = group)) +
  geom_bar(stat = "identity") +
  labs(title = "Total Sales by Tone", 
       x = "Tone", 
       y = "Total Sales") +
  scale_x_discrete(labels = c("Group A" = "Conversational",
                              "Group B" = "Alarming")) +
  scale_fill_discrete(labels = c("Group A" = "Conversational",
                                 "Group B" = "Alarming"))
```