---
title: "Project2 Technical Notebook"
output: html_notebook
---
load in tidyverse and  libraries needed for statistical tests
```{r}
library(tidyverse)
library(pwr)
library(car)
library(effectsize)
```

read in the email metrics csv 
```{r}
email_csv <- read_csv("data/email_metrics.csv")

```
read in the last email before order csv 
```{r}
last_email_csv <- read_csv("data/last_email_before_order.csv")
```
we can see that the columns of group and group code are NOT the same thing; group specifies the general category (conversational approach vs direct) whereas group_code has subcategories within each approach. 0 2 and 4 likely correspond to the days of the week (mon, wed, fri) the emails were sent 

```{r}
distinct(email_csv, group, group_code)
```
create copy of original email mentrics dataset  to preserve original version
```{r}
cleaned_email_csv <- email_csv 

```

do the same for last_email_sent dataset 

```{r}
cleaned_last_email_csv <- last_email_csv
```

based on findings, due to the column number_ordered not having a single instance of 0 orders placed, we can interpret NA as there being 0 orders 

```{r}
cleaned_email_csv$number_orders[is.na(cleaned_email_csv$number_orders)] <- 0
```

we can remove the ...1 column as it is just the row index 

```{r}
cleaned_email_csv <- cleaned_email_csv %>% 
  select(-`...1`)
```

```{r}
cleaned_last_email_csv <- cleaned_last_email_csv %>%
  select(-`...1`)
```

ensure that there isnt a instance of there being 0 emails sent - there is for number_opens although we can leave this as NaN because 0 opens means the email itself was never interacted with by a user 
```{r}
any(cleaned_email_csv$emails_sent == 0)
any(cleaned_email_csv$number_opens == 0)
```
create calculated field columns that measures open rate, click rate, open to click rate, unsub rate, and order rate 

```{r}
cleaned_email_csv <- cleaned_email_csv %>% 
  mutate(open_rate = (number_opens / emails_sent),
         click_rate = (number_clicks / emails_sent),
         open_to_click_rate = (number_clicks / number_opens),
         unsubscribe_rate = (number_unsub / emails_sent),
         order_rate = (number_orders / emails_sent)
  )
         

```


calculate the mean value based on tone for open rate (conversational vs alarming) 
```{r}

cleaned_email_csv %>% 
  group_by(group) %>% 
  summarise(mean_val = mean(open_rate, na.rm = TRUE))

```

running the leven's test based on tone  and the open rate, we recieve a p value of  0.0015. our signfiicance rate is 0.05. since it indicates unequal variances between group A and group B (p = 0.0015), equal variance can't be assumed 
```{r}
leveneTest(open_rate ~ group, data = cleaned_email_csv)
```
because open rate is continuous and we're comparing two independent groups (group a vs group b), we run a two sample t-test. levene's test was performed to assess the difference in variance of each group, and since our p value was below 0.05, welch's  t-test is used 

```{r}
t.test(open_rate ~ group, alternative = "two.sided", conf.level = 0.95, data = cleaned_email_csv)
```

create a sample of data, group by tone, and categorize two variables of the avg open rate and standard deviation for open rate 
```{r}
openRate <- cleaned_email_csv %>% 
  group_by(group) %>% 
  summarise(AvgOpens = mean(open_rate), SdOpens = sd(open_rate))
```

plot tone on x axis, avg open rate as the y axis, and differentiate by tone. then i  created two summary statistics of both the avg open rate, and standard deviation of open rate.  I used a column graph to present my findings along with a error bar with the minimum and maximum intervals. each error bar represent +-1 standard deviation from the mean, showing variability of open rates between group A and group B. 

```{r}
ggplot(openRate, aes(x = group, 
                     y = AvgOpens,
                     fill = group)) +
                    labs(title = "Average Open Rate of Emails Based on Tone ",
                       x = "Tone",
                       y = "Open Rate",
                       fill = "Tone") +
  geom_col() + 
  geom_errorbar(aes(ymin = AvgOpens - SdOpens, ymax = AvgOpens + SdOpens)) + scale_x_discrete(labels = c("Group A" = "Conversational", "Group B" = "Alarming")) + scale_fill_discrete(labels = c("Group A" = "Conversational", "Group B" = "Alarming"))
```


calculate the mean value for click rate based on tone (conversational vs alarming)
```{r}
cleaned_email_csv %>% 
  group_by(group) %>%
  summarise(mean_val = mean(click_rate, na.rm = TRUE))
```

regarding  click rate, our levenes test returns a p value of 0.66, since it is greater than the significance level (0.05), we dont have sufficient evidence to conclude that the variances between group A and group B are different

```{r}
leveneTest(click_rate ~ group, data = cleaned_email_csv)
```
since the p value from the leven's test is greater than the significance level of 0.05, we can assume equal variance between group A and group B. therefore ,a independent two sample t test is used rather than a welch's t test. We specify it to be two-sided as we are testing for any difference between group A and group B rather than asserting a specific group has a higher amount 

```{r}
t.test(click_rate ~ group, alternative = "two.sided", conf.level = 0.95, var.equal = TRUE, data = cleaned_email_csv) 
```
a boxplot was used to visualize distribution of the click rate across group A and group B, to showcase the spread and outliers between conversational and alarming tones 

```{r}
ggplot(cleaned_email_csv, aes(x = group, y = click_rate)) +
  geom_boxplot() + 
  labs(title = "Rate of Clicks by Conversational Tone",
       x = "Tone",
       y = "Click Rate"
      ) +
  scale_x_discrete(labels = c("Group A" = "Conversational", "Group B" = "Alarming")) + 
  scale_fill_discrete(labels = c("Group A" = "Conversational", "Group B" = "Alarming"))
```





calculate the mean value for click to open rate based on tone (conversational vs alarming) 
```{r}
cleaned_email_csv %>% 
  group_by(group) %>%
  summarise(mean_val = mean(open_to_click_rate, na.rm = TRUE))


```
regarding open to click rate, we recieved a p value of 0.71. since our p value is greater than our significance level 0.71 > 0.05, we assume there is equal variance between group A and group B 
```{r}
leveneTest(open_to_click_rate ~ group, data = cleaned_email_csv)
```
 we can run a standard two sample t test due to the results from our leven's test with our p value being greater than our significance level, so we use equal variances along wtih a two sided approach. 

```{r}
t.test(open_to_click_rate ~ group, alternative = "two.sided", conf.level = 0.95, var.equal = TRUE, data = cleaned_email_csv) 
```

aggregate  the data into a sample dataframe that will be plotted. I filtered out any NaN values from the calculated column. 

```{r}
open_to_click_rate_graph <- cleaned_email_csv %>%
  filter(!is.nan(open_to_click_rate)) %>% 
  group_by(group) %>% 
  summarise(AvgOpenClickRate = mean(open_to_click_rate), StdOpenClickRate = sd(open_to_click_rate)) 
```

plot tone on x axis, avg open to click rate  as the y axis, and differentiate by tone. then created two summary statistics of both the avg open to click rate, and standard deviation of open to click rate.  I used a column graph to present  findings along with a error bar with the minimum and maximum intervals, showing variability between group A and group B . 


```{r}
ggplot(open_to_click_rate_graph, aes(x = group, y = AvgOpenClickRate, fill = group)) + 
  labs(title = "Average Open to Click Rate of Emails Based on Tone",
       x = "Tone",
       y = "Open to Click Rate",
       fill = "Tone") + 
  geom_col() + 
  geom_errorbar(aes(ymin = AvgOpenClickRate - StdOpenClickRate, ymax = AvgOpenClickRate + StdOpenClickRate)) + 
  scale_x_discrete(labels = c("Group A" = "Conversational", "Group B" = "Alarming")) +
  scale_fill_discrete(labels = c("Group A" = "Conversational", "Group B" = "Alarming"))

```



