---
title: "Project2 Technical Notebook"
output: html_notebook
---
load in tidyverse library 
```{r}
library(tidyverse)
```
read in the email metrics csv 
```{r}
email_csv <- read_csv("data/email_metrics.csv")

```
read in the last email before order csv 
```{r}
last_email_csv <- read_csv("data/last_email_before_order.csv")
```
we can see that the columns of group and group code are NOT the same thing; group specifies the general category (conversational approach vs direct) whereas group_code has subcategories within each approach. 0 2 and 4 likely correspond to the days of the week (mon, wed, fri) the emails were sent 

```{r}
distinct(email_csv, group, group_code)
```
create copy of original email mentrics dataset  to preserve original version
```{r}
cleaned_email_csv <- email_csv 

```

do the same for last_email_sent dataset 

```{r}
cleaned_last_email_csv <- last_email_csv
```

based on findings, due to the column number_ordered not having a single instance of 0 orders placed, we can interpret NA as there being 0 orders 

```{r}
cleaned_email_csv$number_orders[is.na(cleaned_email_csv$number_orders)] <- 0
```

we can remove the ...1 column as it is just the row index 

```{r}
cleaned_email_csv <- cleaned_email_csv %>% 
  select(-`...1`)
```

```{r}
cleaned_last_email_csv <- cleaned_last_email_csv %>%
  select(-`...1`)
```

ensure that there isnt a instance of there being 0 emails sent - there is for number_opens although we can leave this as NaN because 0 opens means the email itself was never interacted with by a user 
```{r}
any(cleaned_email_csv$emails_sent == 0)
any(cleaned_email_csv$number_opens == 0)
```
create calculated field columns that measures open rate, click rate, open to click rate, unsub rate, and order rate 

```{r}
cleaned_email_csv <- cleaned_email_csv %>% 
  mutate(open_rate = (number_opens / emails_sent),
         click_rate = (number_clicks / emails_sent),
         open_to_click_rate = (number_clicks / number_opens),
         unsubscribe_rate = (number_unsub / emails_sent),
         order_rate = (number_orders / emails_sent)
  )
         

```


calculate the mean value based on tone for open rate (conversational vs alarming) 
```{r}

cleaned_email_csv %>% 
  group_by(group) %>% 
  summarise(mean_val = mean(open_rate, na.rm = TRUE))

```


calculate the mean value for click rate based on tone (conversational vs alarming)
```{r}
cleaned_email_csv %>% 
  group_by(group) %>%
  summarise(mean_val = mean(click_rate, na.rm = TRUE))
```


calculate the mean value for click to open rate based on tone (conversational vs alarming) 
```{r}
cleaned_email_csv %>% 
  group_by(group) %>%
  summarise(mean_val = mean(open_to_click_rate, na.rm = TRUE))


```


