---
title: "Project2 Technical Notebook"
output: html_notebook
---
load in tidyverse and  libraries needed for statistical tests
```{r}
library(tidyverse)
library(pwr)
library(car)
library(effectsize)
```

read in the email metrics csv 
```{r}
email_csv <- read_csv("data/email_metrics.csv")

```
read in the last email before order csv 
```{r}
last_email_csv <- read_csv("data/last_email_before_order.csv")
```
we can see that the columns of group and group code are NOT the same thing; group specifies the general category (conversational approach vs direct) whereas group_code has subcategories within each approach. 0 2 and 4 likely correspond to the days of the week (mon, wed, fri) the emails were sent 

```{r}
distinct(email_csv, group, group_code)
```
create copy of original email mentrics dataset  to preserve original version
```{r}
cleaned_email_csv <- email_csv 

```

do the same for last_email_sent dataset 

```{r}
cleaned_last_email_csv <- last_email_csv
```

based on findings, due to the column number_ordered not having a single instance of 0 orders placed, we can interpret NA as there being 0 orders 

```{r}
cleaned_email_csv$number_orders[is.na(cleaned_email_csv$number_orders)] <- 0
```

we can remove the ...1 column as it is just the row index 

```{r}
cleaned_email_csv <- cleaned_email_csv %>% 
  select(-`...1`)
```

```{r}
cleaned_last_email_csv <- cleaned_last_email_csv %>%
  select(-`...1`)
```

ensure that there isnt a instance of there being 0 emails sent - there is for number_opens although we can leave this as NaN because 0 opens means the email itself was never interacted with by a user 
```{r}
any(cleaned_email_csv$emails_sent == 0)
any(cleaned_email_csv$number_opens == 0)
```
create calculated field columns that measures open rate, click rate, open to click rate, unsub rate, and order rate 

```{r}
cleaned_email_csv <- cleaned_email_csv %>% 
  mutate(open_rate = (number_opens / emails_sent),
         click_rate = (number_clicks / emails_sent),
         open_to_click_rate = (number_clicks / number_opens),
         unsubscribe_rate = (number_unsub / emails_sent),
         order_rate = (number_orders / emails_sent)
  )
         

```


calculate the mean value based on tone for open rate (conversational vs alarming) 
```{r}

cleaned_email_csv %>% 
  group_by(group) %>% 
  summarise(mean_val = mean(open_rate, na.rm = TRUE))

```

running the leven's test based on tone  and the open rate, we recieve a p value of  0.0015. our signfiicance rate is 0.05. As 0.0015 < 0.05, we can assume that there is enough evidence that the group variances between group A and B are different and we reject the null hypothesis.
```{r}
leveneTest(open_rate ~ group, data = cleaned_email_csv)
```
due to the results of  our p value being smaller than our significance level, we run a welch test 
```{r}
t.test(open_rate ~ group, alternative = "two.sided", conf.level = 0.95, data = cleaned_email_csv)
```
create sample of data, group by tone, and categorize two variables of the avg open rate and standard deviation for open rate 
```{r}
openRate <- cleaned_email_csv %>% 
  group_by(group) %>% 
  summarise(AvgOpens = mean(open_rate), SdOpens = sd(open_rate))
```

plot tone on x axis, avg open rate as the y axis, and differentiate by tone. then create two summary statistics of both the avg open rate, and standard deviation of open rate.  I used a column graph to present my findings along with a error bar with the minimum and maximum intervals representing the areas where I expect the majority of my data to fall in. For a conversational tone, our min interval is approximately  ~0.024, and our max interval is approximately  ~ 0.14. For a alarming tone our min interval is similarly ~0.024, although the max interval is much higher, being ~0.24

```{r}
ggplot(openRate, aes(x = group, 
                     y = AvgOpens,
                     fill = group)) +
                    labs(title = "Average Open Rate of Emails Based on Tone ",
                       x = "Tone",
                       y = "Open Rate",
                       fill = "Tone") +
  geom_col() + 
  geom_errorbar(aes(ymin = AvgOpens - SdOpens, ymax = AvgOpens + SdOpens)) + scale_x_discrete(labels = c("Group A" = "Conversational", "Group B" = "Alarming")) + scale_fill_discrete(labels = c("Group A" = "Conversational", "Group B" = "Alarming"))
```


calculate the mean value for click rate based on tone (conversational vs alarming)
```{r}
cleaned_email_csv %>% 
  group_by(group) %>%
  summarise(mean_val = mean(click_rate, na.rm = TRUE))
```

regarding  click rate, we recieved a p value of 0.66. since our p value is greater than our significance level 0.66 > 0.05,  we do not have enough evidence to conclude there is signficiant variance between the two groups 

```{r}
leveneTest(click_rate ~ group, data = cleaned_email_csv)
```


```{r}
t.test(click_rate ~ group, alternative = "two.sided", conf.level = 0.95, var.equal = TRUE, data = cleaned_email_csv) 
```

```{r}
ggplot(cleaned_email_csv, aes(x = group, y = click_rate)) +
  geom_boxplot() + 
  labs(title = "Rate of Clicks by Conversational Tone",
       x = "Tone",
       y = "Click Rate"
      ) +
  scale_x_discrete(labels = c("Group A" = "Conversational", "Group B" = "Alarming")) + 
  scale_fill_discrete(labels = c("Group A" = "Conversational", "Group B" = "Alarming"))
```





calculate the mean value for click to open rate based on tone (conversational vs alarming) 
```{r}
cleaned_email_csv %>% 
  group_by(group) %>%
  summarise(mean_val = mean(open_to_click_rate, na.rm = TRUE))


```
regarding open to click rate, we recieved a p value of 0.71. since our p value is greater than our significance level 0.71 > 0.05, there is not enough evidence to claim that there is variance between the two groups

```{r}
leveneTest(open_to_click_rate ~ group, data = cleaned_email_csv)
```
```{r}
t.test(open_to_click_rate ~ group, alternative = "two.sided", conf.level = 0.95, var.equal = TRUE, data = cleaned_email_csv) 
```
```{r}
open_to_click_rate_graph <- cleaned_email_csv %>%
  filter(!is.nan(open_to_click_rate)) %>% 
  group_by(group) %>% 
  summarise(AvgOpenClickRate = mean(open_to_click_rate), StdOpenClickRate = sd(open_to_click_rate)) 
```


```{r}
ggplot(open_to_click_rate_graph, aes(x = group, y = AvgOpenClickRate, fill = group)) + 
  labs(title = "Average Open to Click Rate of Emails Based on Tone",
       x = "Tone",
       y = "Open to Click Rate",
       fill = "Tone") + 
  geom_col() + 
  geom_errorbar(aes(ymin = AvgOpenClickRate - StdOpenClickRate, ymax = AvgOpenClickRate + StdOpenClickRate)) + 
  scale_x_discrete(labels = c("Group A" = "Conversational", "Group B" = "Alarming")) +
  scale_fill_discrete(labels = c("Group A" = "Conversational", "Group B" = "Alarming"))

```

```{r}
library(tidyverse)

email_metrics <- read_csv("data/email_metrics.csv")

email_clean_day <- email_metrics %>% 
  filter(date_email_sent >= as.Date("2004-03-01"),
         date_email_sent <= as.Date("2004-06-30")) %>%
  mutate(
    unsubscribe_rate = number_unsub / emails_sent,
    order_rate = number_orders / emails_sent,
    day_name = case_when(
      day_of_week == 0 ~ "Monday",
      day_of_week == 2 ~ "Wednesday",
      day_of_week == 4 ~ "Friday",
      TRUE ~ NA_character_
    ),
    day_name = factor(day_name, levels = c("Monday", "Wednesday", "Friday"))
  )
#Graph 1: Unsubscribe Rate by Day
unsub_summary <- email_clean_day %>% 
  drop_na(day_name, unsubscribe_rate) %>%
  group_by(day_name) %>%
  summarise(mean_unsub = mean(unsubscribe_rate), .groups = "drop")

ggplot(unsub_summary, aes(x = day_name, y = mean_unsub, fill = day_name)) +
  geom_col() + geom_text(aes(label = mean_unsub), vjust = 1.5, color = "black") +
  labs(title = "Unsubscribe Rate by Day of Week (Mar-Jun 2004)",
       x = "Day of Week",
       y = "Average Unsubscribe Rate")

unsub_anova <- aov(unsubscribe_rate ~ day_name, 
                   data = email_clean_day %>% 
                     drop_na(day_name, unsubscribe_rate))
summary(unsub_anova)

#Graph 2: Order Rate by Day
order_summary <- email_clean_day %>%
  drop_na(day_name, order_rate) %>% 
  group_by(day_name) %>% 
  summarise(mean_order_rate = mean(order_rate), .groups = "drop")

ggplot(order_summary, aes(x = day_name, y = mean_order_rate, fill = day_name)) +
  geom_col() + geom_text(aes(label = mean_order_rate), vjust = 1.5, color = "black") +
  labs(title = "Order Rate by Day of Week (Mar-Jun 2004)",
       x = "Day of Week",
       y = "Average Order Rate")

order_anova <- aov(order_rate ~ day_name, 
                   data = email_clean_day %>% drop_na(day_name, order_rate))
summary(order_anova)

#Joining data
last_email <- read_csv("data/last_email_before_order.csv")
sales_data_csv <- read_csv("data/cleaned_sales_data.csv")

sales_clean <- sales_data_csv %>%
  select(ORDERNUMBER, SALES) %>%
  drop_na(ORDERNUMBER, SALES)

orders_clean_day <- last_email %>%
  inner_join(sales_clean, by = "ORDERNUMBER") %>%
  filter(date_email_sent >= as.Date("2004-03-01"),
         date_email_sent <= as.Date("2004-06-30")) %>%
  mutate(
    day_name = case_when(
      day_of_week == 0 ~ "Monday",
      day_of_week == 2 ~ "Wednesday",
      day_of_week == 4 ~ "Friday",
      TRUE ~ NA_character_
    ),
    day_name = factor(day_name, levels = c("Monday", "Wednesday", "Friday"))
  ) %>% 
  drop_na(day_name, SALES)

#Graph 3: Sales Amount by Day
sales_summary <- orders_clean_day %>% 
  group_by(day_name) %>% 
  summarise(mean_sales = mean(SALES), .groups = "drop")

ggplot(sales_summary, aes(x = day_name, y = mean_sales, fill = day_name)) +
  geom_col() + geom_text(aes(label = mean_sales), vjust = 1.5, color = "black") +
  labs(title = "Average Sales Amount by Day of Week (Mar-Jun 2004)",
       x = "Day of Week",
       y = "Average Sales Amount")

sales_anova <- aov(SALES ~ day_name, data = orders_clean_day)
summary(sales_anova)

summary(unsub_anova)
summary(order_anova)
summary(sales_anova)
```
A one-way ANOVA test was conducted to determine whether the email send days (Mon, Wed, Fri) significantly affected unsubscribe rate, order rate, and sales amount during March - June 2004. For unsubscribe rate, the results were not statistically significant, F(2, 75) = 0.626, p = 0.538, which indicates or shows that there's no meaningful difference in unsubscribe behavior across the three send/push days and it barely reaches the significance level. While yes it is true the bar chart shows that Wednesday has a slightly higher unsubscribe rate compared to Monday and Friday, the one-way ANOVA (Analysis of Variance) indicates that this different is not, stats-wise at least, significant. This could be random variation in unsubscribe rates across the days. 

For order rate, the ANOVA approached significance but did not meet the 0.05 threshold, F(2, 18) = 0.715, p = 0.503. Although Friday appears to generate a slightly greater order rate and Wednesday the lowest, the difference is not, stats-wise at least, significant.

For sales amount, the ANOVA showed little to no significant difference between the days as F(2, 191) = 0.152, p = 0.859, suggesting the average revenue per order does not vary meaningfully based on whether emails were pushed on Mondays, Wednesdays, or Fridays. Overall, at the 0.05 significance level, there isn't a statistical evidence that email send days significantly causes an impact on the unsubscribe rate, order rate, or sales amount during this period. 
