---
title: "Project2 Technical Notebook"
output: html_notebook
---
load in tidyverse and  libraries needed for statistical tests
```{r}
library(tidyverse)
library(pwr)
library(car)
library(effectsize)
```

read in the email metrics csv 
```{r}
email_csv <- read_csv("data/email_metrics.csv")

```
read in the last email before order csv 
```{r}
last_email_csv <- read_csv("data/last_email_before_order.csv")
```
we can see that the columns of group and group code are NOT the same thing; group specifies the general category (conversational approach vs direct) whereas group_code has subcategories within each approach. 0 2 and 4 likely correspond to the days of the week (mon, wed, fri) the emails were sent 

```{r}
distinct(email_csv, group, group_code)
```
create copy of original email mentrics dataset  to preserve original version
```{r}
cleaned_email_csv <- email_csv 

```

do the same for last_email_sent dataset 

```{r}
cleaned_last_email_csv <- last_email_csv
```

based on findings, due to the column number_ordered not having a single instance of 0 orders placed, we can interpret NA as there being 0 orders 

```{r}
cleaned_email_csv$number_orders[is.na(cleaned_email_csv$number_orders)] <- 0
```

we can remove the ...1 column as it is just the row index 

```{r}
cleaned_email_csv <- cleaned_email_csv %>% 
  select(-`...1`)
```

```{r}
cleaned_last_email_csv <- cleaned_last_email_csv %>%
  select(-`...1`)
```

ensure that there isnt a instance of there being 0 emails sent - there is for number_opens although we can leave this as NaN because 0 opens means the email itself was never interacted with by a user 
```{r}
any(cleaned_email_csv$emails_sent == 0)
any(cleaned_email_csv$number_opens == 0)
```
create calculated field columns that measures open rate, click rate, open to click rate, unsub rate, and order rate 

```{r}
cleaned_email_csv <- cleaned_email_csv %>% 
  mutate(open_rate = (number_opens / emails_sent),
         click_rate = (number_clicks / emails_sent),
         open_to_click_rate = (number_clicks / number_opens),
         unsubscribe_rate = (number_unsub / emails_sent),
         order_rate = (number_orders / emails_sent)
  )
         

```


calculate the mean value based on tone for open rate (conversational vs alarming) 
```{r}

cleaned_email_csv %>% 
  group_by(group) %>% 
  summarise(mean_val = mean(open_rate, na.rm = TRUE))

```

running the leven's test based on tone  and the open rate, we recieve a p value of  0.0015. our signfiicance rate is 0.05. As 0.0015 < 0.05, we can assume that there is enough evidence that the group variances between group A and B are different and we reject the null hypothesis.
```{r}
leveneTest(open_rate ~ group, data = cleaned_email_csv)
```
due to the results of  our p value being smaller than our significance level, we run a welch test 
```{r}
t.test(open_rate ~ group, alternative = "two.sided", conf.level = 0.95, data = cleaned_email_csv)
```
create sample of data, group by tone, and categorize two variables of the avg open rate and standard deviation for open rate 
```{r}
openRate <- cleaned_email_csv %>% 
  group_by(group) %>% 
  summarise(AvgOpens = mean(open_rate), SdOpens = sd(open_rate))
```

plot tone on x axis, avg open rate as the y axis, and differentiate by tone. then create two summary statistics of both the avg open rate, and standard deviation of open rate.  I used a column graph to present my findings along with a error bar with the minimum and maximum intervals representing the areas where I expect the majority of my data to fall in. For a conversational tone, our min interval is approximately  ~0.024, and our max interval is approximately  ~ 0.14. For a alarming tone our min interval is similarly ~0.024, although the max interval is much higher, being ~0.24

```{r}
ggplot(openRate, aes(x = group, 
                     y = AvgOpens,
                     fill = group)) +
                    labs(title = "Average Open Rate of Emails Based on Tone ",
                       x = "Tone",
                       y = "Open Rate",
                       fill = "Tone") +
  geom_col() + 
  geom_errorbar(aes(ymin = AvgOpens - SdOpens, ymax = AvgOpens + SdOpens)) + scale_x_discrete(labels = c("Group A" = "Conversational", "Group B" = "Alarming")) + scale_fill_discrete(labels = c("Group A" = "Conversational", "Group B" = "Alarming"))
```


calculate the mean value for click rate based on tone (conversational vs alarming)
```{r}
cleaned_email_csv %>% 
  group_by(group) %>%
  summarise(mean_val = mean(click_rate, na.rm = TRUE))
```

regarding  click rate, we recieved a p value of 0.66. since our p value is greater than our significance level 0.66 > 0.05,  we do not have enough evidence to conclude there is signficiant variance between the two groups 

```{r}
leveneTest(click_rate ~ group, data = cleaned_email_csv)
```


```{r}
t.test(click_rate ~ group, alternative = "two.sided", conf.level = 0.95, var.equal = TRUE, data = cleaned_email_csv) 
```

```{r}
ggplot(cleaned_email_csv, aes(x = group, y = click_rate)) +
  geom_boxplot() + 
  labs(title = "Rate of Clicks by Conversational Tone",
       x = "Tone",
       y = "Click Rate"
      ) +
  scale_x_discrete(labels = c("Group A" = "Conversational", "Group B" = "Alarming")) + 
  scale_fill_discrete(labels = c("Group A" = "Conversational", "Group B" = "Alarming"))
```





calculate the mean value for click to open rate based on tone (conversational vs alarming) 
```{r}
cleaned_email_csv %>% 
  group_by(group) %>%
  summarise(mean_val = mean(open_to_click_rate, na.rm = TRUE))


```
regarding open to click rate, we recieved a p value of 0.71. since our p value is greater than our significance level 0.71 > 0.05, there is not enough evidence to claim that there is variance between the two groups

```{r}
leveneTest(open_to_click_rate ~ group, data = cleaned_email_csv)
```
```{r}
t.test(open_to_click_rate ~ group, alternative = "two.sided", conf.level = 0.95, var.equal = TRUE, data = cleaned_email_csv) 
```
```{r}
open_to_click_rate_graph <- cleaned_email_csv %>%
  filter(!is.nan(open_to_click_rate)) %>% 
  group_by(group) %>% 
  summarise(AvgOpenClickRate = mean(open_to_click_rate), StdOpenClickRate = sd(open_to_click_rate)) 
```


```{r}
ggplot(open_to_click_rate_graph, aes(x = group, y = AvgOpenClickRate, fill = group)) + 
  labs(title = "Average Open to Click Rate of Emails Based on Tone",
       x = "Tone",
       y = "Open to Click Rate",
       fill = "Tone") + 
  geom_col() + 
  geom_errorbar(aes(ymin = AvgOpenClickRate - StdOpenClickRate, ymax = AvgOpenClickRate + StdOpenClickRate)) + 
  scale_x_discrete(labels = c("Group A" = "Conversational", "Group B" = "Alarming")) +
  scale_fill_discrete(labels = c("Group A" = "Conversational", "Group B" = "Alarming"))

```
```{r}
# Abdir's section
library(tidyverse)
email_metrics <- read_csv("data/email_metrics.csv")
names(email_metrics)
spec(email_metrics)
day_data <- email_metrics %>% 
  mutate(send_date = as.Date(date_email_sent))
day_unsub <- day_data %>% 
  mutate(unsubscribe_rate = number_unsub / emails_sent,
         day_of_week = factor(day_of_week, 
                              levels = c(0, 2, 4),
                              labels = c("Mon", "Wed", "Fri"))
         ) %>%
  filter(
    send_date >= as.Date("2004-03-01"),
           send_date <= as.Date("2004-06-30")
    )

unsub_summary <- day_unsub %>%
  group_by(day_of_week) %>%
  summarise(mean_unsub = mean(unsubscribe_rate, na.rm = TRUE), .groups = "drop")

ggplot(unsub_summary, aes(x = factor(day_of_week), y = mean_unsub)) +
  geom_col(fill = "cyan") +
  labs(title = "Unsubscribe Rate by Day of Week",
       x = "Day of Week",
       y = "Average Unsubscribe Rate"
       )

unsub_anova <- aov(unsubscribe_rate ~ factor(day_of_week), data = day_unsub)
summary(unsub_anova)
```
```{r}
library(tidyverse)
email_metrics <- read_csv("data/email_metrics.csv")
day_data_filtered <- email_metrics %>% 
  mutate(send_date = as.Date(date_email_sent),
         order_rate = number_orders / emails_sent,
         day_of_week = factor(day_of_week, 
                              levels = c(0, 2, 4),
                              labels = c("Mon", "Wed", "Fri")))
order_summary <- day_data %>%
  filter(
    send_date >= as.Date("2004-03-01"),
           send_date <= as.Date("2004-06-30")
    )

order_summary <- day_data_filtered %>%
  group_by(day_of_week) %>%
  summarise(mean_order_rate = mean(order_rate, na.rm = TRUE),
            .groups = "drop")

ggplot(order_summary, aes(x = factor(day_of_week), y = mean_order_rate, na.rm = TRUE)) +
  geom_col(fill = "cyan") +
  labs(title = "Order Rate by Day of Week",
       x = "Day of Week",
       y = "Average Order Rate")

order_anova <- aov(order_rate ~ factor(day_of_week), data = day_data_filtered)
summary(order_anova)
```
```{r}
library(tidyverse)
last_email <- read_csv("data/last_email_before_order.csv")
sales_data_csv <- read_csv("data/sales_data_sample.csv")

# Joining sales with last_email using ORDERNUMBER
orders_df <- last_email %>%
  inner_join(
    sales_data_csv %>% select(ORDERNUMBER, SALES),
    by = "ORDERNUMBER"
  ) %>%
  mutate(
    day_of_week = factor(day_of_week,
                         levels = c(0, 2, 4),
                         labels = c("Mon", "Wed", "Fri"))
  ) %>%
  filter(date_email_sent >= as.Date("2004-03-01"),
         date_email_sent <= as.Date("2004-06-30")
         )

sales_summary <- orders_df %>% 
  group_by(day_of_week) %>%
  summarise(mean_sales = mean(SALES, na.rm = TRUE),
            .groups = "drop")

ggplot(sales_summary, aes(x = day_of_week, y = mean_sales)) +
  geom_col(fill = "cyan") +
  labs(title = "Average Sales Amount by Day of Week (Mar-Jun 2004)",
       x = "Day of Week",
       y = "Average Sales Amount")

sales_anova <- aov(SALES ~ day_of_week, data = orders_df)
summary(sales_anova)
```
```{r}
summary(order_anova)
summary(unsub_anova)
summary(sales_anova)
```



