---
title: "Project2 Technical Notebook"
output: html_notebook
---
load in tidyverse and  libraries needed for statistical tests
```{r}
library(tidyverse)
library(pwr)
library(car)
library(effectsize)
```

read in the email metrics csv 
```{r}
email_csv <- read_csv("data/email_metrics.csv")

```
read in the last email before order csv 
```{r}
last_email_csv <- read_csv("data/last_email_before_order.csv")
```
we can see that the columns of group and group code are NOT the same thing; group specifies the dasdas category (conversational approach vs direct) whereas group_code has subcategories within each approach. 0 2 and 4 likely correspond to the days of the week (mon, wed, fri) the emails were sent 

```{r}
distinct(email_csv, group, group_code)
```
create copy of original email mentrics dataset  to preserve original version
```{r}
cleaned_email_csv <- email_csv 

```

do the same for last_email_sent dataset 

```{r}
cleaned_last_email_csv <- last_email_csv
```

based on findings, due to the column number_ordered not having a single instance of 0 orders placed, we can interpret NA as there being 0 orders 

```{r}
cleaned_email_csv$number_orders[is.na(cleaned_email_csv$number_orders)] <- 0
```

we can remove the ...1 column as it is just the row index 

```{r}
cleaned_email_csv <- cleaned_email_csv %>% 
  select(-`...1`)
```

```{r}
cleaned_last_email_csv <- cleaned_last_email_csv %>%
  select(-`...1`)
```

ensure that there isnt a instance of there being 0 emails sent - there is for number_opens although we can leave this as NaN because 0 opens means the email itself was never interacted with by a user 
```{r}
any(cleaned_email_csv$emails_sent == 0)
any(cleaned_email_csv$number_opens == 0)
```
create calculated field columns that measures open rate, click rate, open to click rate, unsub rate, and order rate 

```{r}
cleaned_email_csv <- cleaned_email_csv %>% 
  mutate(open_rate = (number_opens / emails_sent),
         click_rate = (number_clicks / emails_sent),
         open_to_click_rate = (number_clicks / number_opens),
         unsubscribe_rate = (number_unsub / emails_sent),
         order_rate = (number_orders / emails_sent)
  )
         

```


calculate the mean value based on tone for open rate (conversational vs alarming) 
```{r}

cleaned_email_csv %>% 
  group_by(group) %>% 
  summarise(mean_val = mean(open_rate, na.rm = TRUE))

```

running the leven's test based on tone  and the open rate, we recieve a p value of  0.0015. our signfiicance rate is 0.05. As 0.0015 < 0.05, we can assume that there is enough evidence that the group variances between group A and B are different and we reject the null hypothesis.
```{r}
leveneTest(open_rate ~ group, data = cleaned_email_csv)
```
due to the results of  our p value being smaller than our significance level, we run a welch test 
```{r}
t.test(open_rate ~ group, alternative = "two.sided", conf.level = 0.95, data = cleaned_email_csv)
```
create sample of data, group by tone, and categorize two variables of the avg open rate and standard deviation for open rate 
```{r}
openRate <- cleaned_email_csv %>% 
  group_by(group) %>% 
  summarise(AvgOpens = mean(open_rate), SdOpens = sd(open_rate))
```

plot tone on x axis, avg open rate as the y axis, and differentiate by tone. then create two summary statistics of both the avg open rate, and standard deviation of open rate.  I used a column graph to present my findings along with a error bar with the minimum and maximum intervals representing the areas where I expect the majority of my data to fall in. For a conversational tone, our min interval is approximately  ~0.024, and our max interval is approximately  ~ 0.14. For a alarming tone our min interval is similarly ~0.024, although the max interval is much higher, being ~0.24

```{r}
ggplot(openRate, aes(x = group, 
                     y = AvgOpens,
                     fill = group)) +
                    labs(title = "Average Open Rate of Emails Based on Tone ",
                       x = "Tone",
                       y = "Open Rate",
                       fill = "Tone") +
  geom_col() + 
  geom_errorbar(aes(ymin = AvgOpens - SdOpens, ymax = AvgOpens + SdOpens)) + scale_x_discrete(labels = c("Group A" = "Conversational", "Group B" = "Alarming")) + scale_fill_discrete(labels = c("Group A" = "Conversational", "Group B" = "Alarming"))
```


calculate the mean value for click rate based on tone (conversational vs alarming)
```{r}
cleaned_email_csv %>% 
  group_by(group) %>%
  summarise(mean_val = mean(click_rate, na.rm = TRUE))
```

regarding  click rate, we recieved a p value of 0.66. since our p value is greater than our significance level 0.66 > 0.05,  we do not have enough evidence to conclude there is signficiant variance between the two groups 

```{r}
leveneTest(click_rate ~ group, data = cleaned_email_csv)
```


```{r}
t.test(click_rate ~ group, alternative = "two.sided", conf.level = 0.95, var.equal = TRUE, data = cleaned_email_csv) 
```

```{r}
ggplot(cleaned_email_csv, aes(x = group, y = click_rate)) +
  geom_boxplot() + 
  labs(title = "Rate of Clicks by Conversational Tone",
       x = "Tone",
       y = "Click Rate"
      ) +
  scale_x_discrete(labels = c("Group A" = "Conversational", "Group B" = "Alarming")) + 
  scale_fill_discrete(labels = c("Group A" = "Conversational", "Group B" = "Alarming"))
```





calculate the mean value for click to open rate based on tone (conversational vs alarming) 
```{r}
cleaned_email_csv %>% 
  group_by(group) %>%
  summarise(mean_val = mean(open_to_click_rate, na.rm = TRUE))


```
regarding open to click rate, we recieved a p value of 0.71. since our p value is greater than our significance level 0.71 > 0.05, there is not enough evidence to claim that there is variance between the two groups

```{r}
leveneTest(open_to_click_rate ~ group, data = cleaned_email_csv)
```
```{r}
t.test(open_to_click_rate ~ group, alternative = "two.sided", conf.level = 0.95, var.equal = TRUE, data = cleaned_email_csv) 
```
```{r}
open_to_click_rate_graph <- cleaned_email_csv %>%
  filter(!is.nan(open_to_click_rate)) %>% 
  group_by(group) %>% 
  summarise(AvgOpenClickRate = mean(open_to_click_rate), StdOpenClickRate = sd(open_to_click_rate)) 
```


```{r}
ggplot(open_to_click_rate_graph, aes(x = group, y = AvgOpenClickRate, fill = group)) + 
  labs(title = "Average Open to Click Rate of Emails Based on Tone",
       x = "Tone",
       y = "Open to Click Rate",
       fill = "Tone") + 
  geom_col() + 
  geom_errorbar(aes(ymin = AvgOpenClickRate - StdOpenClickRate, ymax = AvgOpenClickRate + StdOpenClickRate)) + 
  scale_x_discrete(labels = c("Group A" = "Conversational", "Group B" = "Alarming")) +
  scale_fill_discrete(labels = c("Group A" = "Conversational", "Group B" = "Alarming"))

```

-----DREW-----



calculate the mean value based on tone for unsubscribe rate (conversational vs alarming)
```{r}
cleaned_email_csv %>% 
  group_by(group) %>% 
  summarise(mean_val = mean(unsubscribe_rate, na.rm = TRUE))
```


check p value
```{r}
leveneTest(unsubscribe_rate ~ group, data = cleaned_email_csv)
```


our p value is < 0.05 so we assume there is evidence to reject the null hypothesis, run a welch test
```{r}
t.test(unsubscribe_rate ~ group, alternative = "two.sided", conf.level = 0.95, data = cleaned_email_csv)
```


create unsubscription graph
```{r}
unsubscribe_rate_graph <- cleaned_email_csv %>%
  group_by(group) %>%
  summarise(
    mean_unsub_rate = mean(unsubscribe_rate, na.rm = TRUE),
  )

ggplot(unsubscribe_rate_graph, aes(x = group, 
                                   y = mean_unsub_rate,
                                   fill = group)) +
  labs(title = "Average Unsubscribe Rate Based on Tone",
       x = "Tone",
       y = "Unsubscribe Rate",
       fill = "Tone") +
  geom_col() +
  scale_x_discrete(labels = c("Group A" = "Conversational",
                              "Group B" = "Alarming")) +
  scale_fill_discrete(labels = c("Group A" = "Conversational",
                                 "Group B" = "Alarming"))
```


calculate the mean value based on tone for order rate (conversational vs alarming)
```{r}
cleaned_email_csv %>% 
  group_by(group) %>% 
  summarise(mean_val = mean(order_rate, na.rm = TRUE))
```


check p value
```{r}
leveneTest(order_rate ~ group, data = cleaned_email_csv)
```


our p value is < 0.05 so we assume there is evidence to reject the null hypothesis, run a welch test
```{r}
t.test(order_rate ~ group, alternative = "two.sided", conf.level = 0.95, data = cleaned_email_csv)
```


make order graph
```{r}
order_rate_graph <- cleaned_email_csv %>%
  group_by(group) %>%
  summarise(
    mean_order_rate = mean(order_rate, na.rm = TRUE),
  )

ggplot(order_rate_graph, aes(x = group, 
                                   y = mean_order_rate,
                                   fill = group)) +
  labs(title = "Average Order Rate Based on Tone",
       x = "Tone",
       y = "Order Rate",
       fill = "Tone") +
  geom_col() +
  scale_x_discrete(labels = c("Group A" = "Conversational",
                              "Group B" = "Alarming")) +
  scale_fill_discrete(labels = c("Group A" = "Conversational",
                                 "Group B" = "Alarming"))
```


create sales csv
```{r}
sales_data_csv <- read_csv("data/sales_data_sample.csv")
```


make orders data frame merge order numbers
```{r}
orders_df <- merge(cleaned_last_email_csv,
                sales_data_csv[, c("ORDERNUMBER", "SALES")], 
                by = "ORDERNUMBER")
head(orders_df, 10)
```


sum sales per order number include group
```{r}
order_number_totals <- aggregate(SALES ~ ORDERNUMBER + group, 
                                 data = orders_df, 
                                 sum)
head(order_number_totals, 45)
```


sum sales per group and find average sales per order number per group
```{r}
group_totals <- aggregate(SALES ~ group, 
                          data = order_number_totals, 
                          sum)
average_sales_group <- aggregate(SALES ~ group, 
                          data = order_number_totals, 
                          mean)
head(group_totals)
head(average_sales_group)
```


find p value
```{r}
leveneTest(SALES ~ group, data = order_number_totals)
```


p value is > 0.05, variance is statistically insignificant, we accept the null hypothesis
```{r}
t.test(SALES ~ group, data = order_number_totals, var.equal = TRUE)
```


sales sum per group (conversational vs alarming)
```{r}
options(scipen = 999)
ggplot(group_totals, aes(x = group, y = SALES, fill = group)) +
  geom_bar(stat = "identity") +
  labs(title = "Total Sales by Tone", 
       x = "Tone", 
       y = "Total Sales") +
  scale_x_discrete(labels = c("Group A" = "Conversational",
                              "Group B" = "Alarming")) +
  scale_fill_discrete(labels = c("Group A" = "Conversational",
                                 "Group B" = "Alarming"))
```


average sales per group (conversational vs alarming)
```{r}
options(scipen = 999)
ggplot(average_sales_group, aes(x = group, y = SALES, fill = group)) +
  geom_bar(stat = "identity") +
  labs(title = "Average Sales by Tone", 
       x = "Tone", 
       y = "Average Sales") +
  scale_x_discrete(labels = c("Group A" = "Conversational",
                              "Group B" = "Alarming")) +
  scale_fill_discrete(labels = c("Group A" = "Conversational",
                                 "Group B" = "Alarming"))
```



-------Drew------

------ Matthew -----

Filter data for the March - June period and calculate the required metrics

```{r}

analysis_data <- cleaned_email_csv %>%
  filter(before_march == FALSE) %>%
  mutate(
    # Map day_of_week to descriptive names
    day_name = case_when(
      day_of_week == 0 ~ "Monday",
      day_of_week == 2 ~ "Wednesday",
      day_of_week == 4 ~ "Friday"
    ),
    day_name = factor(day_name, levels = c("Monday", "Wednesday", "Friday"))
  )
```

--- DAY VS OPEN RATE ---

Calculate mean open rate by day

```{r}
analysis_data %>% 
  group_by(day_name) %>% 
  summarise(mean_val = mean(open_rate, na.rm = TRUE))
```

Check for variance equality (Levene's Test)

```{r}
leveneTest(open_rate ~ day_name, data = analysis_data)
```

Statistical Test: Welch's ANOVA used because variances are unequal (p < 0.05)

```{r}
oneway.test(open_rate ~ day_name, data = analysis_data, var.equal = FALSE)
```
High f value and p < 0.05 means there is significant chance the day affects the open rate.

Prepare data for Column Graph

```{r}
openRateDay <- analysis_data %>%
  group_by(day_name) %>%
  summarise(AvgOpens = mean(open_rate), SdOpens = sd(open_rate))
```

Visualization: Column Chart
```{r}

ggplot(openRateDay, aes(x = day_name, y = AvgOpens, fill = day_name)) +
  geom_col() +
  geom_errorbar(aes(ymin = AvgOpens - SdOpens, ymax = AvgOpens + SdOpens), width = 0.2) +
  labs(title = "Average Open Rate by Day of Week",
       x = "Day of Week",
       y = "Open Rate",
       fill = "Day") +
  theme_minimal()
```
--- DAY VS OPEN RATE ANALYSIS ---

Monday is the leader for initial engagement with an average open rate significantly higher than Wednesday or Friday. Because the variance in open rates was unequal across days I used Welch's ANOVA and confirmed a significant relationship p < 0.05 between the day of the week and the likelihood of a recipient opening an email. The large error bar for Monday indicates higher volatility, but the overall rate is still higher than either Wednesday or Friday.


---DAY VS CLICK RATE---

```{r}
analysis_data %>% 
  group_by(day_name) %>% 
  summarise(mean_click_rate = mean(click_rate, na.rm = TRUE))
```
Check for variance equality (Levene's Test)

```{r}
leveneTest(click_rate ~ day_name, data = analysis_data)
```
Statistical Test: Standard ANOVA used because variances were not unequal (p  > 0.5)

```{r}
click_anova <- aov(click_rate ~ day_name, data = analysis_data)
summary(click_anova)
```
Visualization: Boxplot (To show distribution consistency)
```{r}

ggplot(analysis_data, aes(x = day_name, y = click_rate, fill = day_name)) +
  geom_boxplot() +
  labs(title = "Click Rate Distribution by Day of Week",
       x = "Day of Week",
       y = "Click Rate") +
  theme_minimal()
```
--- DAY VS CLICK RATE ANALYSIS ---

The average click rate appears more uniform across the three days than the open rate. A standard ANOVA was used here because variances were equal p > 0.5 in Levene's Test, showing that the day of the week has a lesser impact on total audience click-through than it does on opens. The box plot shows that while Monday has a higher ceiling for click rate, the median click rates for Monday, Wednesday, and Friday are relatively the same.

--- DAY VS OPEN TO CLICK RATE ---

Calculate otc mean values by day
```{r}
analysis_data %>% 
  group_by(day_name) %>% 
  summarise(mean_otc_rate = mean(open_to_click_rate, na.rm = TRUE))
```

Check for variance equality (Levene's Test)
```{r}
leveneTest(open_to_click_rate ~ day_name, data = analysis_data)
```
Statistical Test: Standard ANOVA used because variances were not unequal (p  > 0.5)
```{r}
otc_anova <- aov(open_to_click_rate ~ day_name, data = analysis_data)
summary(otc_anova)
```
Prepare data for Column Graph
```{r}
otc_graph_data <- analysis_data %>%
  filter(!is.nan(open_to_click_rate)) %>% 
  group_by(day_name) %>%
  summarise(AvgOTC = mean(open_to_click_rate), SdOTC = sd(open_to_click_rate))
```

Visualization: Column Chart with Error Bars
```{r}

ggplot(otc_graph_data, aes(x = day_name, y = AvgOTC, fill = day_name)) +
  geom_col() +
  geom_errorbar(aes(ymin = AvgOTC - SdOTC, ymax = AvgOTC + SdOTC), width = 0.2) +
  labs(title = "Average Open to Click Rate by Day of Week",
       x = "Day of Week",
       y = "Open to Click Rate",
       fill = "Day") +
  theme_minimal()
```
--- DAY VS OPEN TO CLICK RATE ANALYSIS ---

Although Monday has more opens, Wednesday is slightly ahead in the Open-to-Click Rate, although it is very slight. With equal variances, the standard ANOVA shows that once an email is opened, the day it was sent has a negligible effect on the recipient's decision to click. So once an email is opened, the click rate is relatively the same, meaning that it is more important that a user opens the email in the first place, because once having done so, the click rate is fairly uniform.


------ Matthew -----